Towards a standardized evaluation of imputation methodology

### 2022-09-22 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
More typos

When algorithms are used the simulator should study the validity of the algorithmic process. Many contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, to generate imputations. As with any iterative algorithm--but especially with imputation algorithms that are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}--algorithmic convergence should be carefully evaluated. \citet{ober21} demonstrate that statistical validity can happen before algorithmic convergence is reached, but this is not a guarantee. Unfortunately, there is no universal quantitative method to diagnose non-convergence in iterative imputation algorithms \citep{zhu15, ober21} and the alternative \citep[visual inspection of the imputation algorithm;][]{buur18} is neither efficient nor fail-proof. As a result, imputation algorithms may be terminated before reaching a stable state, which could yield sub-optimal imputations and under-estimated performance of the method. Problems with producing stable imputations are not exclusive to iterative imputation algorithms. Any method may run into failures of the imputation-generating process and subsequently lack results, for example due to over-parameterization errors.

Every imputation workflow should therefore contain an evaluation of the obtained imputations. Even though inspecting each imputation may be labor-intensive due to the number of imputations generated in a simulation study, we highly recommend simulators to consider the following aspects:

>>>>

When algorithms are used the simulator should study the validity of the algorithmic process. Many contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, to generate imputations. As with any iterative algorithm--but especially with imputation algorithms that are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}--algorithmic convergence should be carefully evaluated. \citet{ober21} demonstrate that statistical validity can happen before algorithmic convergence is reached, but this is not a guarantee. Unfortunately, there is no universal quantitative method to diagnose non-convergence in iterative imputation algorithms \citep{zhu15, ober21} and the alternative \citep[visual inspection of the imputation algorithm;][]{buur18} is neither efficient nor fail-proof. As a result, imputation algorithms may be terminated before reaching a stable state, which could yield sub-optimal imputations and under-estimated performance of the method. Problems with producing stable imputations are not exclusive to iterative imputation algorithms. Any method may run into failures of the imputation-generating process and subsequently lack results, for example due to over-parameterisation errors.

Every imputation workflow should therefore contain an evaluation of the obtained imputations. Even though inspecting each imputation may be labour-intensive due to the number of imputations generated in a simulation study, we highly recommend simulators consider the following aspects:
